Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2019-03-04 paper/paper.tex
minor tweaks to text

\subsection*{A system to identify and enrich for rare IFN+ cells}

%%% start IFNrare figure %%%
\begin{figure}
\centerline{
{\bf \Large A}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
{\bf Reporter cells to identify and enrich infections that activate IFN expression.}
{\bf (A)} The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\SUPPFILE{reporter_sequences}).
In A549 cells, the reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\SUPPFIG{reportervalidation}A).
{\bf (B)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the type III IFN reporter (see \SUPPFIG{IFNrare_flow} for full flow cytometry data).
The plot also shows uninfected cells, and cells infected with Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}
\end{figure}
%%% end IFNrare figure %%%

A challenge in studying IFN induction by influenza virus is the rareness with which infection triggers its expression.
Studies using reporters~\cite{killip2017single} and our own previous single-cell transcriptomics~\cite{russell2018extreme} have found that IFN induction is rare in cell lines infected with influenza.
IFN expression also appears to be rare among cells in influenza-infected mice~\cite{kallfass2013visualizing}, a conclusion supported by our re-analysis of single-cell transcriptomics of influenza-infected mice by Steuerman et al~\cite{steuerman2018dissection} (\SUPPFIG{mice}).

To efficiently identify and enrich rare IFN+ cells, we created A549 cells that carried IFN reporters consisting of a type I (\textit{IFNB1}) or type III (\textit{IFNL1}) promoter driving expression of a cell-surface protein (LNGFR$\Delta$C~\cite{bonini1997hsv,ruggieri1997cell}) followed by a fluorescent protein (\FIG{IFNrare}A).
Cells that activate the IFN reporter can be enriched by magnetic-activated cell sorting (MACS) or identified by flow cytometry.
The reporters were efficiently activated by a strain of Sendai virus~\cite{strahle2006sendai} that potently induces IFN (\SUPPFIG{reportervalidation}A), and activation of the type I and type III IFN reporters was highly correlated in our cells (\SUPPFIG{reportervalidation}B; further validated by the single-cell transcriptomics below).
Therefore, for the rest of this paper, we use ``IFN expression'' to refer to combined expression of type I and III IFNs.

We generated a stock of ``wild-type'' A/WSN/1933 (H1N1) influenza (hereafter referred to as ``WSN''), and found that it activated the IFN reporter in $\sim$0.5\% of infected cells  (\FIG{IFNrare}B), a frequency roughly comparable to that in our prior single-cell transcriptomics of influenza-infected cells.

>>>>

\subsection*{A system to identify and enrich rare IFN+ cells}

%%% start IFNrare figure %%%
\begin{figure}
\centerline{
{\bf \Large A}\includegraphics[width=0.35\textwidth,valign=t]{figures/IFN_stochastic/IFN_reporter/IFN_reporter.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.23\textwidth,valign=t]{figures/IFN_stochastic/Flow/ifn_percent.pdf}
}
\caption{
{\bf Reporter cells to identify and enrich infections that activate IFN expression.}
{\bf (A)} The reporter consists of an IFN promoter that drives expression of a cell-surface protein amenable to MACS and a fluorescent protein.
We created reporters with type I and type III IFN promoters (\SUPPFILE{reporter_sequences}).
In A549 cells, the reporters were efficiently activated by an IFN-inducing strain of Sendai virus (\SUPPFIG{reportervalidation}A).
{\bf (B)}
Frequency of IFN induction upon infection with the influenza virus stock used in the single-cell studies in this paper, as quantified using the type III IFN reporter (see \SUPPFIG{IFNrare_flow} for full flow cytometry data).
The plot also shows uninfected cells, and cells infected with saturating amounts of Sendai virus.
The limit of detection of 0.05\% is indicated with a dashed line, and numbers show the median of three measurements.
}
\label{fig:IFNrare}
\end{figure}
%%% end IFNrare figure %%%

A challenge in studying IFN induction by influenza virus is its rareness at the level of single cells.
Studies using reporters~\cite{killip2017single} and our own previous single-cell transcriptomics~\cite{russell2018extreme} have found that IFN induction occurs at a frequency of $<$1\% in cell lines infected with influenza.
IFN expression also appears to be rare among cells in influenza-infected mice~\cite{kallfass2013visualizing}, a conclusion supported by our re-analysis of single-cell transcriptomics of influenza-infected mice by Steuerman et al~\cite{steuerman2018dissection} (\SUPPFIG{mice}).

To efficiently identify and enrich rare IFN+ cells, we created A549 cells that carried IFN reporters consisting of a type I (\textit{IFNB1}) or type III (\textit{IFNL1}) promoter driving expression of a cell-surface protein (LNGFR$\Delta$C~\cite{bonini1997hsv,ruggieri1997cell}) followed by a fluorescent protein (\FIG{IFNrare}A).
Cells that activate the IFN reporter can be enriched by magnetic-activated cell sorting (MACS) or identified by flow cytometry.
The reporters were efficiently activated by infection with saturating amounts of a strain of Sendai virus~\cite{strahle2006sendai} that potently induces IFN (\SUPPFIG{reportervalidation}A), and activation of the type I and type III IFN reporters was highly correlated in our cells (\SUPPFIG{reportervalidation}B; further validated by the single-cell transcriptomics below).
For the rest of this paper, we use ``IFN expression'' to refer to combined expression of type I and III IFNs.

We generated a stock of A/WSN/1933 (H1N1) influenza (hereafter referred to as ``WSN''), and found that it activated the IFN reporter in $\sim$0.5\% of infected cells  (\FIG{IFNrare}B), a frequency roughly comparable to that in our prior single-cell transcriptomics of influenza-infected cells.
